---
title: "Assignement1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assignement1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(Ass1Caba0009)
```

# Introduction
This is the vignette detailing the requirements for assignment 1.

# Chapter 1
This chapter deals with basic metrics between multivariable data including sample means, sample variances, and covariances.

## Problem 1.1
In this problem we were asked to determine the sample means, sample variances, and covariance between two vectors of length 7. The vectors used and metrics calculated are provided below.
```{r, echo=FALSE}
x1 <- c(3,4,2,6,8,2,5)
x2 <- c(5,5.5,4,7,10,5,7.5)
mean_var_cov(x1,x2)
```

## Problem 1.2
In this problem we were provided data relating used car sales by age of the car in years and selling price in thousands of dollars.

### Part A
In part A we were asked to construct a scatter plot of the data as well as including the marginal dot diagrams.
```{r, echo=FALSE, fig.width=6, fig.height=6, fig.align='center'}
x1 <- c(1,2,3,3,4,5,6,8,9,11)
x2 <- c(18.95,19.00,17.95,15.54,14.00,12.95,8.94,7.49,6.00,3.99)
scatter_marginal(x1,x2,"Age of Car (years)","Price of Car (thousands of dollars)")
```
This plot shows the distribution of data and a general trend towards older cars being cheaper. The problem asked for dot diagrams. However, after trying for about an hour to create separate plots in ggplot2 and mending them together with cowplot, the result was not near as informative as the graph above. Therefore, I included this graph in lieu of presenting the other more poorly formatted option.

### Part B
This question asks us to infer the sign of the covariance $s_{12}$ from the scatter plot. Given the negative correlation between car age and selling price (i.e., as age increase, selling price decreases) the covariance between the two variables can be inferred to be negative.

### Part C
This part asked for the sample means, sample variances, sample covariance, and sample correlation coefficient. The function from problem 1 "mean_var_cov" can be used here as well as their is a conditional for displaying the correlation coefficient.
```{r, echo=FALSE}
x1 <- c(1,2,3,3,4,5,6,8,9,11)
x2 <- c(18.95,19.00,17.95,15.54,14.00,12.95,8.94,7.49,6.00,3.99)
mean_var_cov(x1,x2,TRUE)
```
The negative correlation coefficient confirms what we inferred in the previous part of this problem.

### Part D
This problem asks us to display the sample mean array, sample variance-covariance array, and the sample correlation array. This is done with a new function that is based on the first mean_var_cov function used in problem 1.1 and problem 1.2c.
```{r, echo=FALSE}
x1 <- c(1,2,3,3,4,5,6,8,9,11)
x2 <- c(18.95,19.00,17.95,15.54,14.00,12.95,8.94,7.49,6.00,3.99)
vec_mean_var_cov(x1,x2)
```

## Problem 1.3
This problem asks us to follow a similar procedure to that outlined above in problem 1.2d, but in this case we will be using three vectors (x1,x2,x3) instead of just two (x1,x2).
```{r, echo=FALSE}
x1 <- c(9,2,6,5,8)
x2 <- c(12,8,6,4,10)
x3 <- c(3,4,0,2,1)
vec_mean_var_cov(x1,x2,x3)
```
Shown above are the sample mean array, sample cov/var matrix, and sample cor matrix as requested by the instructions in problem 1.3.

## Problem 1.4
For this problem a data file was provided containing information about the world's 10 largest companies with regard to their sales (x1), profits (x2), and assets (x3). The first step is to import the data and then we will work to construct scatter plots with marginal diagrams similar to problem 1.2 for the x1 and x2 variables. Data was imported using the RStudio 'import dataset' tool, col names where given, and then the command usethis::use_data(Problem_1_4) was called to create an .rda object in the data folder which could then be properly documented in R/data.R.  

### Part A
In part A, we were asked to create a scatter plot of the first two columns in the table (Sales,Profits) with marginal diagrams. We can reuse the same function from problem 1.2 here after making modifications to allow different axis labels.
```{r, echo=FALSE, fig.width=6, fig.height=6, fig.align='center'}
sales <- Problem_1_4[,1]
profits <- Problem_1_4[,2]
scatter_marginal(sales,profits,"Sales (billions)","Profits (billions)")
```
The graph above shows the relationship between sales and profits wherein increasing sales leads to increasing profits.

### Part B
We were asked to find the sample means, sample variances, sample covariance, and sample correlation coefficient. This can be accomplished with the 'mean_var_cov' function.
```{r, echo=FALSE}
sales <- Problem_1_4[,1]
profits <- Problem_1_4[,2]
mean_var_cov(sales,profits,cor=TRUE)
```
The positive correlation coefficient allows us to conclude that as sales (x1) increased so does profits (x2).

## Problem 1.5
This problem uses the same data as problem 1.4 but looks at the relationships between profits and assets (x2,x3) as well as sales and assets (x1,x3)

### Part A
In this part we were asked to create two scatter plots: one for profits vs assets (x2,x3) and one for sales vs assets (x1,x3). This can be accomplished with our scatter_marginal function from earlier.
```{r, echo=FALSE, fig.width=6, fig.height=6, fig.align='center'}
sales <- Problem_1_4[,1]
profits <- Problem_1_4[,2]
assets <- Problem_1_4[,3]
scatter_marginal(profits,assets,"Profits (billions)","Assets (billions)")
```

The plot of profits vs assets seems to display a slight negative correlation. However, given the sparsity of the data and the rather random distribution, this conclusion is not very strong.

```{r, echo=FALSE, fig.width=6, fig.height=6, fig.align='center'}
scatter_marginal(sales,assets,"Sales (billions)","Assets (billions)")
```

Th plot of sales vs assets also seems to show a slight negative correlation. But, once again due to the sparsity of the data, it is difficult to draw any concrete conclusions based purely on a scatter plot.

### Part B
We were now asked to compute the column means, variance/covariance matrix, and the correlation coefficient matrix involving all three variables: sales, profits, and assets (x1,x2,x3).
```{r, echo=FALSE}
vec_mean_var_cov(sales,profits,assets)
```

These metrics confirm our suspicions from the scatter plots showing that there is a negative correlation between profits and assets (x2,x3) and well as for sales and assets (x1,x3).

## Problem 1.6
In this problem we were asked to create marginal diagrams and find some metrics for a data frame containing seven variables which is stored in the data/ folder as 'Table_1_5'. In order to accomplish this the data was imported similarly to before using the built in rstudio function as well as the usethis library.

### Part A
In order to plot the all of the different combinations of variables we will utilize a new function called subplots_scatter_marginal which relies on the original scatter_marginal function to create the plots and then arranges them.
```{r, echo=FALSE, fig.width=8, fig.height=20, fig.align='center'}
data <- Table_1_5
subplots_scatter_marginal(data)
```

These plots allow us a brief glimpse of the potential relationships between all 7 of the presented variables

### Part B
Now we were asked to display the sample mean vector, the covariance matrix, and the correlation matrix. This can be done by using a new function called df_metrics which provides these statistics from a dataframe input.
```{r, echo=FALSE}
df_metrics(data)
```

From the correlation coefficient plot we can see that there is a negative relationship between wind and all variables except for hydrocarbons (HC) which has a positive correlation. Solar radiation carries a positive correlation with carbon monoxide (CO), nitrogen dioxide (NO2), ozone (O3), and hydrocarbons (HC). Solar radiation carries a negative correlation with wind and nitric oxide (NO). Carbon monoxide (CO) possesses a positive correlation with all varialbes except wind. Nitric oxide (NO) carries a positive correlation with carbon monoxide (CO), nitrogen dioxide (NO2), and hydrocarbons (HC) with the others bearing a negative correlation. Nitrogen dioxide (NO2) shows a positive correlation with all variables except for wind. Ozone (O3) displays a negative correlation with wind and nitric oxide (NO) but a positive correlation with all other variables. Finally, hydrocarbons (HC) show a positive correlation with all other variables.

# Chapter 2
This chapter deals with matrix and vector manipulation and relationships.

## Problem 2.1
This problem presents two vectors $\textbf{x'} = [5,1,3], \textbf{y'} = [-1,3,1]$

### Part A
Part A asks us to simply graph the two vectors. In order to do this we will call a function called graph_vectors.
```{r, echo=FALSE, fig.width=6, fig.height=6, fig.align='center'}
vec_x <- c(5,1,3)
vec_y <- c(-1,3,1)
graph_vectors(vec_x,vec_y)
```
This plot shows the 3D representation of $\textbf{x'}$ and $\textbf{y'}$

### Part B
In this part we were asked to find the length of $\textbf{x}$, the angle between $\textbf{x}$ and $\textbf{y}$ as well as the projection of $\textbf{y}$ on $\textbf{x}$. To do this we will make use of some simple functions included in R as well as user built functions for the angle between vectors and the projection of one vector on another.
```{r, echo=FALSE}
x_length <- sqrt(sum(vec_x^2))
cat("Vector x has length:", x_length, "\n")
angle_list <- angle_vectors(vec_x,vec_y)
cat("Angle between vectors in degrees", angle_list$Degrees, "\n")
cat("Angle between vectors in radians", angle_list$Radians, "\n")
projection_y_x <- round(projection_vectors(vec_x,vec_y), 4)
cat("Projection of y on x [", projection_y_x, "]")
```

### Part C
Part C asks us to plot the mean corrected vectors. We will utilize the plotting function from earlier after converting the vectors.
```{r, echo=FALSE, fig.width=6, fig.height=6, fig.align='center'}
mean_x <- mean(vec_x)
mean_y <- mean(vec_y)
mean_corr_vec_x <- vec_x - mean_x
mean_corr_vec_y <- vec_y - mean_y
cat("Mean corrected vector x [", mean_corr_vec_x, "]\n")
cat("Mean corrected vector y [", mean_corr_vec_y, "]\n")
graph_vectors(mean_corr_vec_x,mean_corr_vec_y)
```

## Problem 2.2
In this problem we are asked to perform several matrix operation including scalar multiplication, matrix multiplication, and matrix transposition. We are provided the following matrices.
$$\textbf{A}=\begin{bmatrix}-1 & 3 \\ 4 & 2\end{bmatrix}, \textbf{B}=\begin{bmatrix}4 & -3 \\ 1 & -2 \\ -2 & 0 \end{bmatrix}, \textbf{C}=\begin{bmatrix}5 \\ -4 \\ 2 \end{bmatrix}$$
```{r, echo=FALSE}
matA <- matrix(c(-1,3,4,2), nrow=2, ncol=2, byrow=TRUE)
matB <- matrix(c(4,-3,1,-2,-2,0), nrow=3, ncol=2, byrow=TRUE)
matC <- matrix(c(5,-4,2), nrow=3, ncol=1, byrow=TRUE)
```


### Part A
In part a we are asked to perform scalar multiplication on matrix A, $5\textbf{A}$
```{r, echo=FALSE}
matA_5 <- matA * 5
cat("Matrix A multiplied by scalar 5\n")
print(matA_5)
```

### Part B
We were asked to multiply matrix B and matrix A, $\textbf{B}\textbf{A}$
```{r, echo=FALSE}
matBA <- matB %*% matA
cat("Matrix B multiplied by matrix A\n")
print(matBA)
```

### Part C
We were asked to multiply the transpose of matrix A and the transpose of matrix B, $\textbf{A'}\textbf{B'}$
```{r, echo=FALSE}
matAtBt <- t(matA) %*% t(matB)
cat("Transpose of A multiplied by transpose of B\n")
print(matAtBt)
```

### Part D
We were asked to multiply the transpose of matrix C by matrix B, $\textbf{C'}\textbf{B}$
```{r, echo=FALSE}
matCtB <- t(matC) %*% matB
cat("Transpose of C multiplied by B\n")
print(matCtB)
```

### Part E
We were asked if the multiplication of matrix A by matrix B is defined, $\textbf{A}\textbf{B}$. In order for matrix multiplication to be defined the number of columns in matrix 1 must be the same as the number of rows for matrix 2. As matrix A has only two columns but matrix B has three rows, the multiplication, $\textbf{A}\textbf{B}$, is not defined.

## Problem 2.3
In this problem we are asked to verify some common matrix properties around transposition with some new matrices defined below.
$$\textbf{A}=\begin{bmatrix}2 & 1 \\ 1 & 3\end{bmatrix}, \textbf{B}=\begin{bmatrix}1 & 4 & 2 \\ 5 & 0 & 3 \end{bmatrix}, \textbf{C}=\begin{bmatrix}1 & 4 \\ 3 & 2 \end{bmatrix}$$
```{r, echo=FALSE}
matA <- matrix(c(2,1,1,3), nrow=2, ncol=2, byrow=TRUE)
matB <- matrix(c(1,4,2,5,0,3), nrow=2, ncol=3, byrow=TRUE)
matC <- matrix(c(1,4,3,2), nrow=2, ncol=2, byrow=TRUE)
```

### Part A
We are asked to confirm that $(\textbf{A}')'=\textbf{A}. To do this we will display the double transpose of matrix A as well as the original matrix A.
```{r, echo=FALSE}
matAtt <- t(t(matA))
cat("Double transpose of A\n")
print(matAtt)
cat("Original matrix A\n")
print(matA)
```
This allows us to confirm that the double transpose of a matrix is equal to the original matrix.

### Part B
We are asked to confirm that the inverse of a transposed matrix is equivalent to the transpose of that same matrix's inverse, $(\textbf{C}')^{-1}=(\textbf{C}^{-1})'$.
```{r, echo=FALSE}
matCti <- solve(t(matC))
matCit <- t(solve(matC))
cat("Inverse of C tranpose")
print(matCti)
cat("Transpose of C inverse")
print(matCit)
```
This confirms that $(\textbf{C}')^{-1}=(\textbf{C}^{-1})'$.

### Part C
We were asked to confirm that the transposition of the product of two matrices is equal to the multiplication in the opposite direction of the separately transposed matrices, $(\textbf{AB})'=\textbf{B}'\textbf{A}'$.
```{r, echo=FALSE}
matAB_t <- t(matA %*% matB)
matBtAt <- t(matB) %*% t(matA)
cat("Transpose of matrix AB")
print(matAB_t)
cat("Transpose B multiplied by transpose A")
print(matBtAt)
```
This confirms that $(\textbf{AB})'=\textbf{B}'\textbf{A}'$.\

### Part D
In the last part of this question we are posed the question if generally for a matrix A of size (mxk) and a matrix B of size (kxl) if $(\textbf{AB})'=\textbf{B}'\textbf{A}'$ is still valid. 

As the limiting factor here is defined matrix multiplication lets take a look at both sides of the equation. On the left side we would have k columns and k rows which is valid for matrix multiplication. On the right side we would have k columns and k rows following transposition which is also valid for matrix multiplication.

## Problem 2.4
In this problem we are required to prove some relations of inverse matrices. We are told that both $\textbf{A}^{-1}$ and $\textbf{B}^{-1}$ exist.

### Part A
Here, we wish to prove $(\textbf{A}')^{-1}=(\textbf{A}^{-1})'$

First we need to write out some definitions.
$$\textbf{AA}^{-1} = \textbf{I}, \textbf{I} = \textbf{I}',(\textbf{AA}^{-1})'=(\textbf{A}^{-1})'\textbf{A}'$$

Now lets start modifying our original statement.
$$(\textbf{A}')^{-1}=(\textbf{A}^{-1})'$$
$$(\textbf{A}')^{-1}\textbf{A}'=(\textbf{A}^{-1})'\textbf{A}'$$
We know that
$$(\textbf{A}^{-1})'\textbf{A}'=\textbf{I}$$
from our definitions above.  

So our equation now becomes
$$(\textbf{A}')^{-1}\textbf{A}'=\textbf{I}$$

Now we can apply a transpose to both sides.
$$((\textbf{A}')^{-1}\textbf{A}')'=\textbf{I}'$$

We know that $(\textbf{A}')'=\textbf{A}$ by the definition of transposition. So our equation becomes
$$\textbf{A}(\textbf{A}^{-1})=\textbf{I}'$$

We can now see that $\textbf{AA}^{-1}=\textbf{I}'=\textbf{I}$ so our equation simplifies to
$$\textbf{I} = \textbf{I}$$  

which is true by the definition of equality, therefore, proving the original statement is also true.

### Part B
We were tasked with proving the following...
$$(\textbf{AB})^{-1}=\textbf{B}^{-1}\textbf{A}^{-1}$$  

We are given that the following is true...
$$(\textbf{B}^{-1}\textbf{A}^{-1})\textbf{AB}=\textbf{B}^{-1}(\textbf{A}^{-1}\textbf{A})\textbf{B}=\textbf{B}^{-1}\textbf{B}=\textbf{I}$$

Let us multiply by sides by $\textbf{AB}$ yielding...
$$(\textbf{AB})^{-1}\textbf{AB}=(\textbf{B}^{-1}\textbf{A}^{-1})\textbf{AB}$$

According to our definitions above we know that the right side is equal to $\textbf{I}$ giving the equation...
$$(\textbf{AB})^{-1}\textbf{AB}=\textbf{I}$$

We also know from the definition of inverse multiplication that the inverse of a matrix multiplied by the original matrix is the identity matrix or $(\textbf{AB})^{-1}\textbf{AB}={I}$ so that leaves us with...
$$\textbf{I}=\textbf{I}$$

This confirms that our original statement, $(\textbf{AB})^{-1}=\textbf{B}^{-1}\textbf{A}^{-1}$ is true.

## Problem 2.5
In this problem we were tasked with determining if a matrix is orthogonal. The matrix is shown below and loaded into the variable matQ.
$$\textbf{Q} = \begin{bmatrix}  \frac{5}{13} & \frac{12}{13} \\ \frac{-12}{13} & \frac{5}{13} \end{bmatrix}$$
```{r, echo=FALSE}
matQ = matrix(c(5/13,12/13,-12/13,5/13), nrow=2, ncol=2, byrow=TRUE)
```

In order to confirm that the matrix is orthogonal we will see if it satisfies the property that states that $\textbf{Q}\textbf{Q}'=\textbf{I}$
```{r, echo=FALSE}
matQt <- t(matQ)
matQQt <- matQ %*% matQt
cat("Result of QQ'")
print(matQQt)
```
As the result is the identity matrix $\textbf{I}$, we know that the matrix $\textbf{Q}$ is orthogonal
